---
title: "bio_class_11"
author: "caroline iacuaniello"
date: "May 9, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### PDB statistics

Import PDB statistics CSV file and calculate percent structures 

```{r}
p <- read.csv("Data Export Summary.csv", row.names = 1)
p
```
```{r}
percent <- (p$Total / sum(p$Total))*100
names(percent) <- row.names(p)
percent
```
```{r}
## Using Bio3D
library(bio3d)

```

Read in HIV-pr structure
```{r}
pdb <- read.pdb("1hsg")
pdb
```
# find the attributes
```{r}
attributes(pdb)
```
```{r}
# look at atoms header
head(pdb$atom)
```
```{r}
# Print a subset of $atom data for the first two atoms
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]
```

```{r}
# Note that individual $atom records can also be accessed like this
pdb$atom$elety[1:2]
```

```{r}
# Which allows us to do the following

plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab="Bfactor")
```
```{r}
str(pdb$atom)
```
# Print a summary of the coordinate data in $xyz
```{r}
pdb$xyz
```
# Examine the row and column dimensions
```{r}
dim(pdb$xyz)
```
# Print coordinates for the first two atom

```{r}
pdb$xyz[ 1, atom2xyz(1:2) ]
```
# Select all C-alpha atoms (return their indices)
```{r}
ca.inds <- atom.select(pdb, "calpha")
ca.inds
```
Q8. Use the Bio3D write.pdb() function to write out a protein only PDB file for viewing in
VMD. Also write out a second separate PDB file for the ligand with residue name MK1
HINT: In Bio3D you can use the trim.pdb() function together with the pub object and your atom
selections such as atom.select(pdb, ’protein’) for the fist file and atom.select(pdb, ‘ligand’) for
the second.
```{r}
inds.ligand <- atom.select(pdb, "ligand")
inds.protein <- atom.select(pdb, "protein")
inds.protein
```

#check what have
```{r}
pdb$atom[inds.ligand$atom,]
```

```{r}
pdb$atom[inds.protein$atom,]
```
# what want is to output new pdb file

```{r}
pdb.ligand <- trim.pdb(pdb, inds=inds.ligand)
pdb.protein <- trim.pdb(pdb, inds=inds.protein)
pdb.ligand
```

```{r}
write.pdb(pdb.ligand, file="1hsg_ligand.pdb")
write.pdb(pdb.protein, file="1hsg_protein.pdb")
```
# Download some example PDB files
```{r}
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)
```
# Extract and align the chains we are interested in
```{r}
pdbs <- pdbaln(files, fit = TRUE)
```
# Print to screen a summary of the 'pdbs' object
```{r}
pdbs

```
Here the returned object is of class pdbs. Note that it contains a xyz numeric matrix of aligned
C-alpha coordinates, a ali matrix of aligned residues, and a resno matrix of aligned residue
numbers (see the list of associated attributes (+ attr)). These attirbutes can be accessed
using the common $ syntax in R. E.g. use pdbs$ali to access the alignment. To access the
first few rows of the alignment matrix we use standard subsetting syntax for matrices in R:

```{r}
# Access the first 5 rows, and 8 columns
pdbs$ali[1:5, 1:8]
```

```{r}
# Associated residues numbers
pdbs$resno[1:5, 1:8]
```

## skip to section 5
5.1 Search and retrieve Adenylate kinase structures
Below we perform a blast search of the PDB database to identify related structures to our query
Adenylate kinase sequence. In this particular example we use function get.seq() to fetch the
query sequence for chain A of the PDB ID 1AKE and use this as input to blast.pdb(). Note
that get.seq() would also allow the corresponding UniProt identifier.

```{r}
# fetch sequence
aa <- get.seq("1ake_A")
aa
```

```{r}
# Blast or hmmer search
b <- blast.pdb(aa)
```
```{r}
# Plot a summary of search results
hits <- plot(b)
# can set cutoff= any number to get more hits
```
```{r}
head(hits$pdb.id)
hits
```
fetch pdbs, and align,
```{r}
# fetch
files <- get.pdb(hits$pdb.id, path = "pdbs", split = TRUE, gzip =
TRUE)
```
```{r}
# Align structures
pdbs <- pdbaln(files)
```
```{r}
# Vector containing PDB codes
ids <- basename.pdb(pdbs$id)
```

```{r}
# Draw schematic alignment
plot(pdbs, labels=ids)
```
```{r}
# find invariant core
core <- core.find(pdbs)
# superimpose all structures to core
pdbs$xyz = pdbfit(pdbs, core)
# Perform PCA
pc.xray <- pca(pdbs)
```

```{r}
# Calculate RMSD
rd <- rmsd(pdbs)
# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)
plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

```{r}
# Visualize first principal component
mktrj(pc.xray, pc=1, file="pc_1.pdb")
```

```{r}
#view in r with new code
# download new

```

