---
title: "Bioinformatics_class_7"
author: "Caroline Iacuaniello"
date: "April 25, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Functions again
We can source any file of R code with the 'source' () function.
```{r}
source("http://tinyurl.com/rescale-R")
```
Lets make sure things are here
```{r}
ls()
```
check rescale function is working

```{r}
rescale(1:10)
```
```{r,eval=FALSE}
rescale( c(1:10, "string"))
```

```{r,eval=FALSE}
rescale2( c(1:10, "string"))
```


## Function for finding missing values in two datasets
Write a 'both_na()' function to do this

```{r}
# Lets define an example x and y
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
is.na(x)
is.na(y)

```

```{r}
which(is.na(x))
```

```{r}
which(!is.na(x))
```

```{r}
(z <- 5)
if(z <- 5) {
  print("more")
} else{print("less")}
```
To count, false = 0, true = 1
```{r}
sum(is.na(x))
```

```{r}
is.na(x)
is.na(y)
```

```{r}
is.na(x) & is.na(y)
```
The & symbol combines

```{r}
sum(is.na(x) & is.na(y))
```

```{r}
which(is.na(x) & is.na(y))
```
My first function can start from this snippet.
```{r}
both_na <- function(x,y) {sum(is.na(x) & is.na(y))}
```
Test it!

```{r}
both_na(x,y)
```
```{r,eval=FALSE}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
both_na(x, y1)
both_na3(x, y2)
```
```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
ans <- both_na3(x,y)
```
```{r}
ans$which
```
# Start with a simple version of the problem
df1 <- data.frame(IDs=c("gene1", "gene2", "gene3"),
 exp=c(2,1,1),
 stringsAsFactors=FALSE)
df2 <- data.frame(IDs=c("gene2", "gene4", "gene3", "gene5"),
 exp=c(-2, NA, 1, 2),
 stringsAsFactors=FALSE)
# Simplify further to single vectors
x <- df1$IDs
y <- df2$IDs
# Now what do we do? Want to find intersection, what is going on in both.
```{r}
x <- df1$IDs
y <- df2$IDs
x
y

```
We can try the 'intersect()' function and the "%in%" functions.

```{r}
intersect(x,y)
```
intersect not very useful, gave answer but no more info

```{r}
x %in% y
```
This is checking if the first thing in x is in y, etc. Try for things in y in x.
```{r}
y %in% x
```
```{r}
x[x %in% y]
y[y %in% x]
```
Now try the cbind() function. binds inputs as columns of new matrix.
These results yield intersects. Use logical output in '%in%' to get at our matching data.

```{r}
x[x %in% y]
y[y %in% x]
```
Lets put these together as columns of a matrix
```{r}
cbind(x[x %in% y],
y[y %in% x])
```
```{r}
cbind( c("Hello", "Help"), c("Please", "me"))
```

```{r}
rbind( c("Hello", "Help"), c("Please", "me"))
```

# Make a gene intersect function
```{r}
gene_intersect <- function(x, y) {
 cbind( x[ x %in% y ], y[ y %in% x ] )
}

#Testing
gene_intersect(x,y)
```
Try with data frame input
```{r}
gene_intersect2(df1,df2)
```

Looks good, this is skateboard :)

```{r}
gene_intersect3(df1, df2)
```
Works, but code is ugly. Make code tighter.
```{r}
gene_intersect4(df1,df2)
```
Test against some data frame with gene 2 found twice 
```{r}
gene_intersect4(df1, df3)
```
GEts a warning message but overall works.

Lets use 'merge()' function for this...

```{r}
merge(df1, df2, by="IDs")
```
```{r}
 source("https://bioconductor.org/biocLite.R")
 biocLite("phyloseq")

```
```{r}
library("phyloseq"); packageVersion("phyloseq")
```

```{r}
# Create a pretend OTU table that you read from a file, called otumat
otumat = matrix(sample(1:100, 100, replace = TRUE), nrow = 10, ncol = 10)
otumat
rownames(otumat) <- paste0("OTU", 1:nrow(otumat))
colnames(otumat) <- paste0("Sample", 1:ncol(otumat))
otumat
taxmat = matrix(sample(letters, 70, replace = TRUE), nrow = nrow(otumat), ncol = 7)
rownames(taxmat) <- rownames(otumat)
colnames(taxmat) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
taxmat
class(otumat)
```
```{r}
class(taxmat)
```

```{r}
library("phyloseq")
OTU = otu_table(otumat, taxa_are_rows = TRUE)
TAX = tax_table(taxmat)
OTU
```
```{r}
TAX
```
```{r}
physeq = phyloseq(OTU, TAX)
physeq
```

```{r}
plot_bar(physeq, fill = "Family")
```

