---
title: "class_13"
author: "caroline iacuaniello"
date: "May 16, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 1000 Genome project data
Read 1000 genome data for MXL dataset
```{r}
genotype<-read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs12936231.csv")
```
```{r}
table(genotype[,2]) / nrow(genotype)*100
```
```{r}
library(seqinr)
library(gtools)
```
### Base quality scores from fastqsanger
```{r}
library(seqinr)
library(gtools)
phred <- asc( s2c("DDDDCDEDCDDDDBBDDDCC@") ) - 33
phred 
```


```{r}
# install.packages("seqinr")

```

```{r}
# install.packages("gtools")
```
## RNA-seq results analysis (interested in assessing genetic differences on a population scale. processed
about ~230 samples and did the normalization on a genome level.  want to find
whether there is any association of the 4 asthma-associated SNPs (rs8067378…) on
ORMDL3 expression. )
```{r}
geno <- read.table("rs8067378_ENSG00000172057.6.txt")
```

```{r}
summary(geno)
```
## we want to know if genotype g/g is different expression than a/a
```{r}
gg <- summary(geno$exp[geno$geno == "G/G"])
ag <- summary(geno$exp[geno$geno == "A/G"])
aa <- summary(geno$exp[geno$geno == "A/A"])

```
## want to make a nice boxplot, my attempt...

```{r}
boxplot(gg)
boxplot(ag)
boxplot(aa)
```
```{r}
boxplot(exp ~ geno, data=geno, notch = TRUE)
```

# how about a nice result? ggplot
```{r}
library(ggplot2)

p <- ggplot(geno, aes(geno,exp)) + geom_boxplot(notch = TRUE)
p
```

